Making all equalities equal

### 2021-09-17 paper.lagda.tex
wrote more about equality instances

\AgdaHide{
\begin{code}
{-# OPTIONS --cubical --type-in-type #-}

module paper where

open import Agda.Primitive.Cubical
\end{code}
}

@@

@@

\begin{code}
module LeibnizEquality {ℓ} {ℓ'} {A : Set ℓ} where
  _≐_ : A → A → Set
  a ≐ b = (P : A → Set ℓ') → P a → P b
\end{code}

@@

\begin{code}
module _ {a} {A : Set a} where
  instance
    ≡-IsEquality : IsEquality {A = A} _≡_
    ≡-IsEquality = eq refl

  instance
    ≡p-IsEquality : IsEquality {A = A} _≡p_
    ≡p-IsEquality = eq (λ {x y} → sym λ i → p-c {x = x} {y = y} i)

module _ {ℓ} where
  univalencePath' : {A B : Type ℓ} → (A ≡ B) ≡ (A ≃ B)
  univalencePath' {A} {B} =
    ua {ℓ-suc ℓ} {A ≡ B} {A ≃ B} (compEquiv (univalence {ℓ} {A} {B})
    (isoToEquiv (iso {ℓ} {ℓ-suc ℓ}
    (λ x → x) (λ x → x) (λ b i → b) λ a i → a)))

  instance
    ≃-IsEquality : IsEquality
      {A = Type ℓ} (λ A B → (A ≃ B))
    ≃-IsEquality = eq λ {A B} → sym
      λ i → univalencePath' {A = A} {B = B} i

liftIso : ∀ {a b} {A : Type a} {B : Type b}
  → Iso {a} {b} A B → Iso {ℓ-max a b} {ℓ-max a b} A B
liftIso {a} f = iso fun inv
  (λ x i → rightInv x i) (λ x i → leftInv x i)

  where open Iso f

open import leibniz
open Leibniz

module FinalEquality {A : Set} where
  open MainResult A

  ≐≅≡ : ∀ {a b} → Iso (a ≐ b) (a ≡p b)
  ≐≅≡ = iso j i (ptoc ∘ ji) (ptoc ∘ ij)

  ≐≡≡' : ∀ {a b} → (a ≐ b) ≡c (a ≡p b)
  ≐≡≡' = let lifted = liftIso ≐≅≡  in isoToPath lifted

  ≐≡≡ : _≐_ {A} ≡c _≡p_
  ≐≡≡ i x y = ≐≡≡' {x} {y} i

  open IsEquality

  instance
    ≐-IsEquality : IsEquality {A = A} _≐_
    ≐-IsEquality = eq λ {x} {y} → ≐≡≡' ∙
      λ i → ≡p-IsEquality {ℓ-zero} .≣-≡-≡ {x} {y} i
\end{code}

>>>>

@@

\subsection{Imports}

First, it will be necessary to give some agda arguments:

\begin{code}
{-# OPTIONS --cubical --cumulativity #-}
module paper where
\end{code}

The cubical flag is necessary because we are using cubical equality,
and the cumulativiy flag is also necessary for level subtyping,

\begin{code}
open import Agda.Primitive.Cubical using (I; i0; i1)
\end{code}

This library loads Cubical Agda Primitives as the equality interval.

@@

\begin{code}
module LeibnizEquality {A : Set} where
  _≐_ : A → A → Set₁
  a ≐ b = (P : A → Set) → P a → P b
\end{code}

@@

The simplest example is the cubical equality hence this equality is already equal to itself.

\begin{code}
module _ {a} {A : Set a} where
  instance
    ≡-IsEquality : IsEquality {A = A} _≡_
    ≡-IsEquality = eq refl
\end{code}

\subsection{Martin-Löf equality}

The proof of Martin-Löf equality is more difficult, but it is already in Cubical library as \AgdaFunction{p-c}.

\begin{code}
  instance
    ≡p-IsEquality : IsEquality {A = A} _≡p_
    ≡p-IsEquality = eq (λ {x y} → sym λ i → p-c {x = x} {y = y} i)
\end{code}

\subsection{Isomorphism}

The isomorphism is an equality between types.

\begin{code}
module _ {ℓ} where
  univalencePath' : {A B : Type ℓ} → (A ≡ B) ≡ (A ≃ B)
  univalencePath' {A} {B} =
    ua {ℓ-suc ℓ} {A ≡ B} {A ≃ B} (compEquiv (univalence {ℓ} {A} {B})
    (isoToEquiv (iso {ℓ} {ℓ-suc ℓ}
    (λ x → x) (λ x → x) (λ b i → b) λ a i → a)))
\end{code}

\AgdaFunction{univalencePath} is already defined in Agda library, but with \AgdaBound{A ≃ B} instead of
\AgdaBound{Lifted (A ≃ B)}. This change can be done because of the cumulativity flag.

\begin{code}
  instance
    ≃-IsEquality : IsEquality
      {A = Type ℓ} _≃_
    ≃-IsEquality = eq λ {A B} → sym
      λ i → univalencePath' {A = A} {B = B} i
\end{code}

\subsection{Leibniz Equality}

The hardest equality to proof that is equalty is the Leibniz Equality.

\begin{code}
liftIso : ∀ {a b} {A : Type a} {B : Type b}
  → Iso {a} {b} A B → Iso {ℓ-max a b} {ℓ-max a b} A B
liftIso {a} f = iso fun inv
  (λ x i → rightInv x i) (λ x i → leftInv x i)
\end{code}

This \AgdaFunction{liftIso} will be used to lift the Isomorphism to types of the same maximum level of both.

\begin{code}
  where open Iso f

open import leibniz
open Leibniz
\end{code}

It is importing the definition of Leibniz equality made by \cite{abel2020leibniz}.
In this work, there is already a proof of the isomorphism between the Leibniz and the Martin-Löf equality.

\begin{code}
module FinalEquality {A : Set} where
  open MainResult A

  ≐≅≡ : ∀ {a b} → Iso (a ≐ b) (a ≡p b)
  ≐≅≡ = iso j i (ptoc ∘ ji) (ptoc ∘ ij)
\end{code}

In Cubical Library, the definition of isomorphism uses cubical equality instead of Martin-Löf equality
when we have to proof that \AgdaBound{∀ x → from (to x) ≡ x} and \AgdaBound{∀ x → to (from x) ≡ x}.
\AgdaFunction{ptoc} is necessary to do this convertion from these equalities.

\begin{code}
  ≐≡≡ : ∀ {a b} → (a ≐ b) ≡c (a ≡p b)
  ≐≡≡ = let lifted = liftIso ≐≅≡  in isoToPath lifted
\end{code}

Using the univalence and \AgdaFunction{liftIso} defined previously, it is possible to transform the isormorphism
into an equality.

\begin{code}
  open IsEquality

  instance
    ≐-IsEquality : IsEquality {A = A} _≐_
    ≐-IsEquality = eq λ {x} {y} → ≐≡≡ ∙
      λ i → ≡p-IsEquality {ℓ-zero} .≣-≡-≡ {x} {y} i
\end{code}

The last pass is to join the three equalities between equalities:
Leibniz to Martin-Löf to cubical equality.
